Improving the communication and transparency of stock assessment using interactive

### 2019-03-06 analysis/paper/manuscript.Rmd
Modify Keith's section and produce the second draft of the manuscript

decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics.  The majority of this sequence of events, sometimes called “the data pipeline”, have not traditionally been part of the peer-review process which sees only the end products.  However, decisions made at these steps increasingly influence the outcome of the study.  Gelman coined the term “the forked garden path” to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path.  Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible, even if it is one’s own analysis.  A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science.  Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work (Lewis et al. 2018).

We submit that the dashboard tool is also a major step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.  Consider the following process.  Through the Canadian Science Advisory Secretariat (CSAS), DFO Science provides peer-reviewed science advice to fisheries managers at regular meetings (Regional Assessment Process).  These meetings are attended by a variety of stake holders including harvesters, academics, and NGOs.  Documents summarizing these processes (research documents, stock assessment reports, and proceedings summaries) are available to the public through the CSAS website.  Ergo the CSAS process strives to uphold the ideal of transparency.  However, as it currently stands, only the end products of an assessment are open and transparent; as with virtually all scientific analyses, the data pipeline of stock assessment is hidden from review.

With the exception of data management, the dashboard removes much this problem – all analyses and diagnostics can be readily reviewed.  In terms of data management, this point is largely ameliorated because the dashboard tool is constructed in Program R using R Markdown and the code for creating the dashboard and all associated analyses are available on Github.  Data are also available through information requests.  So the entire data pipeline is transparent as well as reproducible, i.e., the requirements of open science are fulfilled.

The dashboard also promotes accessibility to the stock assessment.  Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted (Banks 20XX). In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model).  However, the dashboard removes many of these difficulties.  The html file contains all of the relevant analyses.  The number of analyses can be expanded upon request and can be distributed prior to meetings and reviewers can assess the results and diagnostics at their leisure.

We submit that the plotly/flex dashboard tool is an important step forward in transparent and reproducible science.  Further, this tool makes stock assessments more accessible to stakeholders.  Beyond stock assessments, it this tool could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.

> **Other draft text**

> A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers. In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009]. Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.

> In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models. The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits. Data presented in tables is incredibly valuable, but as human beings, seeing patterns can be challenging. Static figures, in contrast, help reveal patterns but they are divorced from the underlying data and, as such, limit detailed explorations. New tools are bridging this gap by making it easier to build interactive figures [@perkel2018].

> **Links to some handy references**

> Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization

> https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/

> https://www.nature.com/articles/s41467-018-03297-7

> https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103

> https://www.nature.com/articles/d41586-018-01322-9

> "The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)

>>>>

The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics. The majority of this sequence of events, sometimes called “the data pipeline”, have not traditionally been part of the peer-review process which sees only the end products. However, decisions made at these steps increasingly influence the outcome of the study. Gelman coined the term “the forked garden path” to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path. Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible, even if it is one’s own analysis. A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science. Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@lewis2018].

We submit that the dashboard tool is also a major step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued. Consider the following process. Through the Canadian Science Advisory Secretariat (CSAS), DFO Science provides peer-reviewed science advice to fisheries managers at regular meetings (Regional Assessment Process). These meetings are attended by a variety of stakeholders including harvesters, academics, and NGOs. Documents summarizing these processes (research documents, stock assessment reports, and proceedings summaries) are available to the public through the CSAS website. Ergo the CSAS process strives to uphold the idea of transparency. However, as it currently stands, only the end products of an assessment are open and transparent; as with virtually all scientific analyses, the data pipeline of stock assessment is hidden from review.

The dashboard removes much of this problem by providing access to all model inputs, outputs and diagnostics in an open and easy to use interactive document. In terms of data management, this point is largely ameliorated because the dashboard tool is constructed in Program R using R Markdown and the code for creating the dashboard and all associated analyses are available on Github. Data are also available through information requests.  So the entire data pipeline is transparent as well as reproducible, i.e., the requirements of open science are fulfilled.

The dashboard also promotes accessibility to the stock assessment. Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted [@banks2011]. In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model). However, the dashboard removes many of these difficulties. The html file contains all of the relevant analyses. The number of analyses can be expanded upon request and can be distributed prior to meetings and reviewers can assess the results and diagnostics at their leisure.

We submit that the dashboard tool is an important step forward in transparent and reproducible science. Further, this tool makes stock assessments more accessible to stakeholders. Beyond stock assessments, interactive documents could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.
